\documentclass[11pt,a4paper]{article}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage[margin=1in]{geometry}
\usepackage{booktabs}
\usepackage{array}
\usepackage{tabularx}
\usepackage{longtable}
\usepackage{listings}
\usepackage{xcolor}
\usepackage{hyperref}
\usepackage{amsmath,amssymb}
\usepackage{enumitem}
\usepackage{float}
\usepackage{caption}

\hypersetup{
    colorlinks=true,
    linkcolor=blue!60!black,
    citecolor=blue!60!black,
    urlcolor=blue!60!black,
}

\lstdefinelanguage{tlaplus}{
    morekeywords={LET,IN,IF,THEN,ELSE,EXCEPT,RECURSIVE,UNCHANGED,CHOOSE,
                  VARIABLES,CONSTANTS,EXTENDS,MODULE,INIT,NEXT,TRUE,FALSE,
                  INVARIANT,INVARIANTS,CONSTANT},
    sensitive=true,
    morecomment=[l]{\\*},
    morestring=[b]",
    literate={|->}{{$\mapsto$}}2
             {\\in}{{$\in$}}1
             {\\A}{{$\forall$}}1
             {\\E}{{$\exists$}}1
             {/\\}{{$\land$}}1
             {\\/}{{$\lor$}}1
             {=>}{{$\Rightarrow$}}1
             {->}{{$\rightarrow$}}1
             {<<}{{$\langle$}}1
             {>>}{{$\rangle$}}1
             {!=}{{$\neq$}}1
             {<=}{{$\leq$}}1
             {>=}{{$\geq$}}1,
}

\lstset{
    basicstyle=\ttfamily\small,
    keywordstyle=\bfseries\color{blue!70!black},
    commentstyle=\itshape\color{gray},
    stringstyle=\color{red!70!black},
    breaklines=true,
    frame=single,
    framerule=0.5pt,
    rulecolor=\color{gray!50},
    xleftmargin=1em,
    framexleftmargin=0.5em,
    numbers=none,
    aboveskip=0.8em,
    belowskip=0.8em,
}

\newcolumntype{L}[1]{>{\raggedright\arraybackslash}p{#1}}
\newcolumntype{C}[1]{>{\centering\arraybackslash}p{#1}}

% Allow more aggressive interword stretching to avoid overfull hboxes
% caused by long \texttt{...} identifiers that cannot be hyphenated.
\emergencystretch=3em

% Permit line breaks inside \texttt after "/" and "_" so that long
% file paths (e.g. MatchingEngine/TheoremsElegant.lean) can wrap.
\makeatletter
\def\BreakableChars#1{\discretionary{}{}{}\ttfamily#1}
\makeatother

\title{From Bounded Model Checking to Mechanized Proof:\\
A Multi-Level Verification Case Study for a\\
Price-Time Priority Matching Engine\thanks{Artifacts available at
\url{https://github.com/formally-verified-exchange/verified-matching-engine}.}}

\author{%
  Ara Aslyan\\
  Strike Technologies, New York\\
  \texttt{aaslyan@striketechnologies.com}
}
\date{04/11/2026}

\begin{document}

\maketitle

\begin{abstract}
We present a multi-level formal verification case study of a
feature-rich price-time priority matching engine. The specification
covers a full order-type suite (LIMIT, MARKET, MARKET-TO-LIMIT,
STOP\_LIMIT, STOP\_MARKET) together with iceberg orders, four
self-trade prevention policies, post-only, Fill-Or-Kill,
minimum-quantity, cancel, and amend semantics. Verification proceeds
at two levels: bounded exhaustive state-space exploration of a TLA+
model, and a mechanized proof in Lean~4 of a Lean reference
implementation that mirrors the same specification.

Within the explored TLC configurations we identify one concrete
specification defect---a FIFO violation caused by triggered stop
orders retaining their original timestamp---and one specification gap
in the interaction between STP DECREMENT and iceberg visibility
management. In Lean, we prove that the main processing theorem
preserves the combined book invariant \texttt{AllInv}, defined as
uncrossed plus sorted on both sides. The proof holds for arbitrary
book sizes with a fuel bound computed from the book state and proved
sufficient rather than assumed. The development consists of a
3{,}986-line constructive proof of the full preservation theorem,
together with a 312-line structural proof of the matching-and-dispose
core via a three-case analysis of the termination condition.

The proof effort itself surfaced a further issue that neither
prose review nor bounded model checking had exposed: the original
arbitrary fuel bound (\texttt{defaultFuel = 100}) made the
preservation theorem false for books with more than a hundred
contra-side orders. We report this as a methodological observation
on the verification pipeline.
\end{abstract}


% ============================================================================
\section{Introduction}
% ============================================================================

Matching-engine correctness is commercially important and surprisingly
difficult to state precisely. Production engines combine a large number of
interacting features---order types, time-in-force policies, self-trade
prevention, hidden quantity, stop triggers, cascading fills---each specified
correctly in isolation, but capable of producing unexpected emergent behavior
when combined. Prose specifications resolve the easy cases and leave the hard
ones to implementor judgment.

Formal methods offer a sharper alternative: write the semantics as an
executable state machine and enumerate its reachable states. This
paper is a case study in doing exactly that for a matching-engine
specification, and then going one step further---lifting the same
semantics into Lean~4 and proving a safety property that bounded
model checking cannot, by itself, establish.

The paper's contributions are:
\begin{enumerate}[nosep]
\item A TLA+ formalization of a feature-rich price-time-priority
      matching-engine specification, together with a suite of
      invariants covering book consistency, queue fairness, and
      order-type safety.
\item A concrete TLC counterexample showing that triggered stop
      orders can violate FIFO if activation does not refresh queue
      priority, and identification of a second defect in the
      interaction between STP DECREMENT and iceberg visibility
      management.
\item A mechanized Lean~4 proof that the processing pipeline
      preserves the combined book invariant \texttt{AllInv}
      (uncrossed plus sorted on both sides) under a minimal
      well-formedness precondition, for arbitrary book sizes, with
      the matching-loop fuel bound computed from the book state and
      proved sufficient rather than assumed.
\item Two complementary proof artifacts with no open obligations:
      a 3{,}986-line constructive proof of the full preservation
      theorem by branch-by-branch induction, and a 312-line
      structural proof of the matching-and-dispose core by
      three-case analysis of the matching termination condition.
      The contrast illustrates two complementary proof styles.
\item A methodological observation from the proof effort itself:
      the original arbitrary fuel bound
      (\texttt{defaultFuel~=~100}) made the preservation theorem
      false for sufficiently populated books.
\item A discussion of specification-guided implementation
      generation, framed as an engineering observation about how a
      precise formal specification constrains downstream coding,
      \emph{not} as a claim of end-to-end refinement between the
      Lean reference and any systems-language implementation.
\end{enumerate}

\subsection{Note on the Development Process}
\label{sec:intro-ai}

The TLA+ model, the Lean~4 proofs, and the C/C++ implementations
discussed in \S\ref{sec:synthesis} were developed interactively with
a large language model acting as a coding assistant. Strategic
decisions---the progress-lemma decomposition, the three-case
analysis in the elegant proof, and the side-parameterization refactor---were human
contributions. The bulk of tactic scripting and operator translation
was produced by the model under that guidance. We mention this
because the per-file size and per-lemma count reported in later
sections reflect that mode of work; the methodology described in
this paper is independent of whether the implementer is a single
engineer, a team, or an AI-augmented workflow.

\subsection{Paper Organization}

The paper is organized as follows.
Section~\ref{sec:semantics} describes the matching-engine semantics
under study. Section~\ref{sec:model} presents the TLA+ model and
the scope of the bounded exploration.
Section~\ref{sec:invariants} enumerates the invariants checked.
Section~\ref{sec:findings} presents the two TLC findings with
counterexample traces, and Section~\ref{sec:fixes} gives the
corresponding specification fixes. Section~\ref{sec:lean} presents
the Lean~4 proof of \texttt{AllInv} preservation: the termination
measure, the progress-lemma decomposition, the mutual induction
across the processing pipeline, and a second, shorter proof via
case analysis of the matching termination condition. An
observation from the proof effort itself---the constant-fuel
soundness issue---is reported in the same section. Section~\ref{sec:synthesis} describes how the
fixed specification was carried into C and C++ implementations and
why we stop short of a refinement claim.
Section~\ref{sec:lessons} collects lessons on feature composition;
Section~\ref{sec:related} positions the work against prior
literature; Section~\ref{sec:limitations} states limitations and
reproducibility conditions; Section~\ref{sec:conclusion} concludes.


% ============================================================================
\section{Matching-Engine Semantics Under Study}
\label{sec:semantics}
% ============================================================================

\subsection{Specification Source}

The object of study is an internal prose specification,
\texttt{matching-engine-formal-spec.md} version~1.2.0, describing a generic
asset-agnostic order matching engine with price-time priority (FIFO). The
specification defines primitive domains, order well-formedness constraints, the
order book structure, a core matching algorithm, post-match disposition, six
order types, four STP policies, iceberg reload semantics, stop triggering with
cascading evaluation, cancel, and amend.

The specification is 900 lines of structured pseudocode. It was written with
care---well-formedness constraints are enumerated, invariants are stated
formally, and edge cases such as MTL price derivation and FOK dry-run
semantics are treated explicitly. It is not a toy specification.


\subsection{Order Types and Features}

The engine supports five order types:
\begin{itemize}[nosep]
\item \textbf{LIMIT}: rests at a specified price; fills at that price or better.
\item \textbf{MARKET}: fills at any available price; never rests on the book.
\item \textbf{MARKET-TO-LIMIT (MTL)}: aggressively matches at any price; locks
      to the first execution price and rests any remainder as a LIMIT order.
\item \textbf{STOP\_LIMIT / STOP\_MARKET}: dormant until a trigger condition on
      the last trade price is satisfied, then converted to LIMIT or MARKET
      respectively and re-entered into the matching pipeline.
\end{itemize}

Order modifiers include:
\begin{itemize}[nosep]
\item \textbf{Iceberg} (\texttt{displayQty}): only
      $\mathit{visibleQty} \leq \mathit{displayQty}$ is shown to the market;
      after the visible slice is consumed, the order reloads with a fresh
      visible slice and loses time priority.
\item \textbf{Post-Only}: the order must not aggress; rejected (under the
      REJECT policy) if it would cross the spread.
\item \textbf{TimeInForce}: GTC, GTD, IOC, FOK, DAY.
\item \textbf{MinQty}: requires a minimum fill on first execution; the
      constraint is cleared after the initial threshold is met, allowing the
      remainder to rest.
\item \textbf{Self-Trade Prevention (STP)}: four policies
      (\textsc{cancel\_newest}, \textsc{cancel\_oldest},
      \textsc{cancel\_both}, \textsc{decrement}) govern what
      happens when an incoming order would trade against a resting
      order in the same STP group. The incoming order's policy
      governs.
\end{itemize}


\subsection{Well-Formedness Constraints}
\label{sec:wf}

The prose specification enumerates 23 well-formedness constraints
(WF-1 through WF-20 plus three a/b suffixed refinements) that define
the \emph{valid inputs} to the engine. Every theorem in the Lean
proof, every TLC invariant, and every rejection message in the
synthesized C++ code traces back to a subset of these constraints.
They are grouped by feature and summarized in
Table~\ref{tab:wf}. The full text with exact quantifiers appears in
\texttt{matching-engine-formal-spec.md~\S2.2}.

\begin{table}[H]
\centering
\caption{Well-formedness constraints (excerpt from the prose
specification). TLA+ \texttt{WellFormed} and Lean
\texttt{Order.wellFormed} model a subset (omitting WF-6/7,
WF-12, WF-17; WF-11 is included but WF-12 is not). Lean's
\texttt{OrderProcOk} requires only the two load-bearing fragments
(WF-13 and its corollary).}
\label{tab:wf}
\small
\begin{tabular}{@{}llL{8cm}@{}}
\toprule
ID & Group & Constraint \\
\midrule
WF-1       & Qty     & $\mathit{quantity} > 0$ \\
WF-2/2a/2b & Price   & \textsc{limit}, \textsc{stop\_limit} require
                       $\mathit{price} > 0$; \textsc{mtl} has $\mathit{price}=\bot$ \\
WF-3       & Price   & \textsc{market}, \textsc{stop\_market} have $\mathit{price}=\bot$ \\
WF-4       & Stop    & Stop orders require $\mathit{stopPrice} > 0$ \\
WF-5       & Stop    & Non-stop orders have $\mathit{stopPrice}=\bot$ \\
WF-6/7     & TIF     & \textsc{gtd} requires $\mathit{expireTime} > \mathit{now}$; otherwise $\mathit{expireTime}=\bot$ \\
WF-8/8a    & TIF     & \textsc{market}$\Rightarrow$\textsc{ioc}/\textsc{fok};
                       \textsc{mtl}$\Rightarrow$\textsc{gtc}/\textsc{day} \\
WF-9/10    & Iceberg & $0 < \mathit{displayQty} \leq \mathit{quantity}$; icebergs must be \textsc{limit} \\
WF-11/12   & Iceberg & Initial $\mathit{remainingQty}=\mathit{quantity}$; $\mathit{visibleQty}=\min(\mathit{displayQty},\mathit{quantity})$ \\
WF-13      & PostOnly& Post-only$\Rightarrow$\textsc{limit} \\
WF-14      & PostOnly& Post-only$\not\Rightarrow$\textsc{ioc}/\textsc{fok} \\
WF-15      & PostOnly& \textsc{market}, \textsc{mtl}$\Rightarrow$$\neg$post-only \\
WF-16      & STP     & $\mathit{selfTradeGroup}=\bot \iff \mathit{stpPolicy}=\bot$ \\
WF-17      & STP     & $\mathit{stpPolicy}\in\{\textsc{cancel\_newest},$ $\textsc{cancel\_oldest},$ $\textsc{cancel\_both},$ $\textsc{decrement}\}$ \\
WF-18      & MinQty  & $0 < \mathit{minQty} \leq \mathit{quantity}$ \\
WF-19      & MinQty  & MinQty$\Rightarrow$$\neg$post-only \\
WF-20      & MinQty  & \textsc{fok}$\Rightarrow$$\mathit{minQty}=\bot$ \\
\bottomrule
\end{tabular}
\end{table}

The Lean predicate \texttt{OrderProcOk} used in the main theorem is
the minimum subset of these constraints that the matching loop
actually depends on. It does not admit every WF-valid order---it
admits every WF-valid order that can enter the matching phase---and
its two conjuncts are refinements of WF-13
(post-only$\Rightarrow$\textsc{limit}, and hence post-only orders
carry a defined price) and a consequence of the same WF-13
(stop orders are not \textsc{limit}$\Rightarrow$$\neg$post-only). The fact that
these two minimal fragments---and no more---suffice was discovered
\emph{during} the proof: earlier attempts omitted them and every
branch of the matching loop that case-split on an undefined price
failed. The proof told us which WF clauses were load-bearing.

\subsection{Core Invariants Stated in the Specification}
\label{sec:inv}

The prose specification states 14 invariants that must hold after
every \texttt{PROCESS} call completes (\S13 of the spec).
Table~\ref{tab:inv-summary} gives the book-structural core; the
complete TLC-checked inventory including safety and progress
properties is given in Table~\ref{tab:invariants}
(\S\ref{sec:invariants}).

\begin{table}[H]
\centering
\caption{Core book invariants (excerpt).}
\label{tab:inv-summary}
\small
\begin{tabular}{@{}llL{7cm}@{}}
\toprule
ID & Group & Statement \\
\midrule
INV-1   & Structural & Every price level has a non-empty order queue \\
INV-2   & Structural & $\forall$ consecutive bid levels $(L_1,L_2)$: $L_1.\mathit{price} > L_2.\mathit{price}$ \\
INV-3   & Structural & $\forall$ consecutive ask levels $(L_1,L_2)$: $L_1.\mathit{price} < L_2.\mathit{price}$ \\
INV-4   & Structural & $\mathit{bestBid} < \mathit{bestAsk}$ or either side empty (\textbf{uncrossed}) \\
INV-5   & Content    & Every resting order has $\mathit{remainingQty} > 0$ \\
INV-6   & Content    & Every resting order has status $\in\{\textsc{new},\textsc{partially\_filled}\}$ \\
INV-7   & Ordering   & FIFO within each price level: consecutive orders have strictly increasing timestamps \\
INV-8   & Content    & No \textsc{market} orders rest on the book \\
INV-9   & Trade      & Every trade executes at the passive order's price \\
INV-10  & Events     & Events emitted in causal order \\
INV-11  & PostOnly   & No trade where the aggressor was post-only \\
INV-12  & STP        & No trade between orders in the same STP group \\
INV-13  & Content    & No \textsc{mtl} orders rest on the book \\
INV-14  & Content    & No $\mathit{minQty}$ is set on any resting order \\
\bottomrule
\end{tabular}
\end{table}

The FIFO invariant INV-7 is central to this paper's first bug. Its
formal statement is:
\begin{quote}
For every price level, consecutive orders satisfy
$o_1.\mathit{timestamp} < o_2.\mathit{timestamp}$.
\end{quote}
This is the formal statement of price-time priority. A violation
means an order with a later arrival time has higher queue position
than one that arrived earlier---the queue is out of order.

\subsection{The \texttt{canMatch} Predicate}
\label{sec:canmatch}

A single predicate governs the matching loop's continuation:

\[
\texttt{canMatchPrice}(o, p) \;\equiv\;
\begin{cases}
\top & o.\mathit{type}\in\{\textsc{market},\textsc{stop\_market}\} \\
o.\mathit{price} \geq p & o.\mathit{side}=\textsc{buy},\; o\text{ priced} \\
o.\mathit{price} \leq p & o.\mathit{side}=\textsc{sell},\; o\text{ priced} \\
\top & o.\mathit{type}=\textsc{mtl},\text{ no trades yet} \\
\text{derived price} & o.\mathit{type}=\textsc{mtl},\text{ after first fill}
\end{cases}
\]

This predicate is the loop guard of \texttt{doMatch}. When the loop
terminates in a non-empty contra side, it terminates \emph{because}
$\neg\texttt{canMatchPrice}(o, \mathit{bestContra})$. This observation
is the seed of the elegant proof in \S\ref{sec:lean-elegant}: if
matching stopped because prices no longer cross, then the book cannot
be crossed after disposition places the residual on the same side as
the aggressor.

\subsection{From Specification to Proof: Traceability}
\label{sec:traceability}

The specification is not documentation---it is the shared vocabulary
that binds TLC and Lean. Every numbered clause appears in at least one
of the two. Table~\ref{tab:trace} makes the chain explicit for the
clauses exercised in this paper.

\begin{table}[H]
\centering
\caption{Spec $\to$ TLA+ $\to$ Lean traceability for the clauses
touched by the main theorem. The Lean column refers to
\texttt{Theorems.lean} or \texttt{Invariants.lean}.}
\label{tab:trace}
\small
\begin{tabular}{@{}lL{3cm}L{3.2cm}L{4.3cm}@{}}
\toprule
Spec clause & TLA+ artefact & Lean artefact & Role \\
\midrule
WF-13       & \texttt{WellFormed} guard & \texttt{OrderProcOk} field 1 & Main-theorem precondition \\
WF-13 (derived) & \texttt{WellFormed} guard & \texttt{OrderProcOk} field 2 & Main-theorem precondition \\
INV-2/3     & not separately checked$^\dagger$ & \texttt{bidsSortedDesc}, \texttt{asksSortedAsc} & Discovered: needed for induction \\
INV-4       & \texttt{BookUncrossed}    & \texttt{BookUncrossed}, part of \texttt{AllInv} & Main-theorem conclusion \\
INV-5       & \texttt{NoGhosts}         & \texttt{NoGhosts} (definition only) & Not proved preserved \\
INV-6       & \texttt{StatusConsistency}& \texttt{statusConsistencyB} (definition only) & Not proved preserved \\
INV-7       & \texttt{FIFOWithinLevel}  & --- & TLC-only \\
INV-8       & \texttt{NoRestingMarkets} & --- & TLC-only \\
INV-11      & \texttt{PostOnlyGuarantee}& --- & TLC-only \\
INV-12      & \texttt{STPGuarantee}     & --- & TLC-only \\
INV-13      & \texttt{NoRestingMTL}     & --- & TLC-only \\
INV-14      & \texttt{NoRestingMinQty}  & \texttt{NoRestingMinQty} & Invariants.lean definition \\
\texttt{canMatch}
            & \texttt{CanMatchPrice}    & \texttt{canMatchPrice}   & Matching loop guard \\
PROCESS \S12& \texttt{ProcessOrder} (11 phases) & \texttt{processOrder} (11 phases) & Mutual induction target \\
\bottomrule
\end{tabular}

\medskip
\raggedright\footnotesize
$^\dagger$Structural sortedness is not stated as a separate TLC
invariant because TLA+ sequences are ordered by construction from a
sorted key. In Lean the sortedness is explicit because book levels are
a \texttt{List PriceLevel} and the induction needs to know both sides
are monotone. This mismatch was discovered \emph{during} the proof:
\texttt{BookUncrossed} alone is not inductive, which forced the
addition of \texttt{AllInv = BookUncrossed $\land$ bidsSortedDesc
$\land$ asksSortedAsc}. The proof refined the invariant.
\end{table}

The lesson is that invariants are \emph{discovered}, not merely
declared. The first row of surprises in Table~\ref{tab:trace}
(sortedness) came from the Lean induction refusing to close;
\texttt{AllInv} is what the proof requires, not what the prose spec
emphasizes. The second row of surprises (\texttt{OrderProcOk} as a
minimal precondition fragment) came from the proof refusing to close
without WF-13 and WF-15. In both cases the proof told us the
specification was under-emphasizing something that was, on reflection,
quite obvious.


% ============================================================================
\section{TLA+ Model and Bounded Exploration}
\label{sec:model}
% ============================================================================

\subsection{Model Structure}

The TLA+ specification (\texttt{MatchingEngine.tla}, approximately 850 lines)
is a direct translation of the prose specification. The state consists of
eight variables:

\begin{lstlisting}[language=tlaplus]
bidQ           : PRICES -> Seq(Order)   (* per-price bid queues *)
askQ           : PRICES -> Seq(Order)   (* per-price ask queues *)
stops          : Set(Order)             (* dormant stop orders *)
lastTradePrice : PRICES \cup {NULL}     (* last execution price *)
postOnlyOK     : Bool                   (* no post-only aggressor trade *)
stpOK          : Bool                   (* no same-group trade *)
nextId         : Nat                    (* next order ID *)
clock          : Nat                    (* logical clock *)
\end{lstlisting}

The model has four actions: \texttt{SubmitOrder}, \texttt{CancelOrder},
\texttt{AmendOrder}, and \texttt{TimeAdvance}. \texttt{SubmitOrder}
nondeterministically selects all valid order parameters and invokes the full
processing pipeline. TLC's exhaustive exploration therefore considers every
valid combination of order type, side, price, quantity, TIF, STP policy,
iceberg flag, post-only flag, and minimum quantity, within the configured
bounds.


\subsection{Processing Pipeline}

The pipeline operator \texttt{ProcessOrder} implements all phases of the prose
specification's \texttt{PROCESS} procedure, including recursive operators for
the matching loop (\texttt{DoMatch}) and stop-trigger cascade
(\texttt{ProcessCascade}, \texttt{ProcessTriggeredStops}). Mutual recursion
between these operators is required to model cascading stop triggers correctly:
a triggered stop may generate trades, which may trigger further stops.

A \texttt{fuel} parameter on \texttt{DoMatch} bounds recursion depth for TLC
termination safety. Under the small quantity bounds used in model checking,
this bound is never reached in practice.


\subsection{Scope of Exploration}

The model is parameterized by \texttt{PRICES}, \texttt{MAX\_QTY},
\texttt{MAX\_ORDERS}, and \texttt{MAX\_CLOCK}. Configurations explored are
shown in Table~\ref{tab:configs}. All configurations include the full
order-type suite, all four STP policies, icebergs, post-only, FOK, MinQty, and
DAY expiry. The ``With Amend'' configuration adds the \texttt{AmendOrder}
action.

\begin{table}[H]
\centering
\caption{Model checking configurations and statistics.}
\label{tab:configs}
\small
\begin{tabular}{@{}lccclrrll@{}}
\toprule
Configuration & Orders & Qty & Prices & Amend & States Gen. & Distinct & Runtime & Result \\
\midrule
Tiny          & 2 & 1 & \{1,2\}   & No  &    2,979,719 &  1,123,333 & 20\,s     & Pass \\
Small         & 2 & 2 & \{1,2\}   & No  &   11,528,343 &  6,037,674 & 1\,m\,42\,s & Pass \\
Medium        & 2 & 2 & \{1,2,3\} & No  &   36,700,016 & 21,261,901 & 5\,m\,34\,s & Pass \\
With Amend    & 2 & 2 & \{1,2\}   & Yes &   25,209,607 &  9,104,902 & 3\,m\,33\,s & Pass \\
3-order       & 3 & 2 & \{1,2\}   & No  & $>$45\,M     & $>$26\,M   & $>$10\,m  & \emph{partial} \\
\bottomrule
\end{tabular}

\medskip
\raggedright\footnotesize
All ``Pass'' results above are post-fix. Prior to fixing the stop-timestamp
bug, the Small configuration produced a counterexample on the
\texttt{FIFOWithinLevel} invariant within the first 11.5 million states. The
3-order run was interrupted before state-space exhaustion; ``partial'' means no
violation was found in the explored states, not clean passage under all
3-order behaviors.
\end{table}


\subsection{What the Model Does Not Cover}

\begin{itemize}[nosep]
\item The model uses a single STP group identifier (\texttt{"G1"}). Multiple
      independent groups are not explored.
\item GTD expiry is modeled via \texttt{TimeAdvance} but without wall-clock
      constraints.
\item Event ordering (INV-10) and market-data publication are not modeled.
\item The model checks specification consistency, not implementation
      correctness.
\end{itemize}


% ============================================================================
\section{Invariants Checked}
\label{sec:invariants}
% ============================================================================

All invariants are checked as TLA+ safety properties (\texttt{INVARIANT}
declarations in the TLC configuration). TLC verifies that each invariant
holds in every reachable state within the explored configurations.

\begin{table}[H]
\centering
\caption{Invariants checked.}
\label{tab:invariants}
\small
\begin{tabular}{@{}llL{4.8cm}ll@{}}
\toprule
ID & TLA+ Operator & Description & TLC & Lean \\
\midrule
INV-1  & \texttt{NoEmptyLevels}    & Every price level has a non-empty order queue & trivial & trivial \\
INV-2  & (sortedness)              & Bids strictly descending & by constr. & $\checkmark$ \\
INV-3  & (sortedness)              & Asks strictly ascending & by constr. & $\checkmark$ \\
INV-4  & \texttt{BookUncrossed}    & $\mathit{bestBid} < \mathit{bestAsk}$ or either side empty & $\checkmark$ & $\checkmark$ \\
INV-5  & \texttt{NoGhosts}         & Every resting order has $\mathit{remainingQty} > 0$ & $\checkmark$ & --- \\
INV-6  & \texttt{StatusConsistency}& Resting orders have status \textsc{new} or \textsc{partial} & $\checkmark$ & --- \\
INV-7  & \texttt{FIFOWithinLevel}  & FIFO within level (strictly increasing timestamps) & fixed$^{\star}$ & --- \\
INV-8  & \texttt{NoRestingMarkets} & \textsc{market} orders never appear on the book & $\checkmark$ & --- \\
INV-9  & (construction)            & Trade price equals the passive order's price & by constr. & by constr. \\
INV-11 & \texttt{PostOnlyGuarantee}& No trade where the aggressor was post-only & $\checkmark$ & --- \\
INV-12 & \texttt{STPGuarantee}     & No trade between orders in the same STP group & $\checkmark$ & --- \\
INV-13 & \texttt{NoRestingMTL}     & \textsc{mtl} orders are converted before resting & $\checkmark$ & --- \\
INV-14 & \texttt{NoRestingMinQty}  & $\mathit{minQty}$ is cleared before an order rests & $\checkmark$ & --- \\
\bottomrule
\end{tabular}

\medskip
\raggedright\footnotesize
IDs match the canonical \S13 summary of the prose specification.
TLC column: $\checkmark$ means bounded exhaustive pass across all
configurations in Table~\ref{tab:configs}. Lean column:
$\checkmark$ means part of \texttt{AllInv} and mechanically proved
for all book sizes by \texttt{process\_preserves\_AllInv}; ---
means not yet lifted to Lean.
$^{\star}$INV-7 was violated under the original specification and
passes after the fix described in \S\ref{sec:bug1}.
\end{table}


% ============================================================================
\section{Findings and Counterexamples}
\label{sec:findings}
% ============================================================================

\subsection{Finding 1: FIFO Violation via Missing Timestamp on Triggered Stops}
\label{sec:bug1}

\paragraph{Severity.} High---directly violates INV-7 (FIFO within price level).

\paragraph{Root cause location.} \S10.1 \texttt{EVALUATE\_STOPS} interaction
with \S6.1 \texttt{INSERT}.

\subsubsection*{The Mechanism}

The prose specification's stop-trigger procedure (\S10.1) converts a dormant
stop order to its base type (LIMIT or MARKET) and re-enters it into the
\texttt{PROCESS} pipeline. The stop order's \texttt{timestamp} field is not
updated at trigger time. It retains the timestamp assigned when the order was
originally submitted and placed into dormancy.

When the triggered LIMIT order does not fully fill and rests on the book via
\texttt{DISPOSE} (\S5.2), it is appended to the back of the appropriate
price-level queue per \S6.1---giving it the lowest priority at that level.
However, \texttt{FIFOWithinLevel} requires that the newly appended order have
a timestamp strictly greater than all orders already in the queue. Because
orders already in the queue arrived after the stop was submitted---they carry
later timestamps---the stop's original timestamp is numerically earlier. The
queue is now out of order.


\subsubsection*{Counterexample Trace}

TLC produced the following minimal counterexample. Times shown are logical
clock values.

\begin{figure}[H]
\centering
\begin{lstlisting}[language={},basicstyle=\ttfamily\footnotesize,frame=single,xleftmargin=0pt,framexleftmargin=0pt]
Step 1: Submit SELL LIMIT  qty=1, price=1      (id=1, ts=1)
        -> Rests on ask:  askQ[1] = [order1(ts=1)]

Step 2: Submit BUY STOP_LIMIT  stop=1, price=1, qty=1  (id=2, ts=2)
        -> lastTradePrice = NULL -> does not trigger
        -> Placed in stops = {order2(ts=2)}

Step 3: Submit BUY LIMIT  qty=2, price=1       (id=3, ts=3)
        -> Matches askQ[1]: fills order1 fully (fillQty=1)
        -> Trade: price=1, aggressor=3, passive=1
        -> lastTradePrice = 1
        -> order3: remainingQty=1, rests on bid
           bidQ[1] = [order3(ts=3)]
        -> Stop cascade: lastTradePrice(1) >= stopPrice(1)
           -> order2 triggers, converts to BUY LIMIT
           -> Timestamp NOT updated: ts=2 (STALE)
           -> Ask side empty, no fill
           -> Dispose: order2 rests on bid
              bidQ[1] = [order3(ts=3), order2(ts=2)]

        FIFOWithinLevel check:
            bidQ[1][1].timestamp = 3
            bidQ[1][2].timestamp = 2
            3 < 2 ?  FALSE  ->  INVARIANT VIOLATED
\end{lstlisting}
\caption{TLC counterexample trace for the FIFO violation. The triggered stop
order (id=2) retains its original timestamp (ts=2) and is appended behind an
order with a later timestamp (ts=3), producing a non-monotonic queue.}
\label{fig:trace}
\end{figure}

\subsubsection*{Why Manual Review Missed This}

The prose specification treats stop triggering and order insertion as separate
sections (\S10.1 and \S6.1 respectively). Each section is locally correct:
\S10.1 correctly describes conversion, and \S6.1 correctly describes
insertion. The interaction---that converted stops carry stale
timestamps---is invisible when reading either section in isolation. A
reviewer must simultaneously hold the conversion path, the insertion path, and
the FIFO invariant in mind to see the problem. TLC holds all three
simultaneously by construction.


\subsection{Finding 2: Iceberg Stranding Under STP DECREMENT}
\label{sec:bug2}

\paragraph{Severity.} Medium---creates a reachable book state with invalid
order geometry.

\paragraph{Root cause location.} \S8.3 DECREMENT case; missing interaction
with \S7.5 iceberg reload.

\subsubsection*{The Mechanism}

The STP DECREMENT policy reduces both the incoming order's
$\mathit{remainingQty}$ and the resting order's $\mathit{visibleQty}$ and
$\mathit{remainingQty}$ by
$\min(\mathit{incoming.remainingQty},\, \mathit{resting.visibleQty})$, without
generating a trade. The iceberg reload logic (\S7.5 step 3) specifies:

\begin{quote}
When $\mathit{visibleQty} = 0$ and $\mathit{remainingQty} > 0$, reload the
visible slice, assign a new timestamp, and move the order to the back of the
queue.
\end{quote}

However, this reload is only specified inside the fill path. The DECREMENT
action (\S8.3) does not reference \S7.5. There is no clause in \S8.3 for the
case where DECREMENT drives $\mathit{visibleQty}$ to zero while
$\mathit{remainingQty} > 0$.

The result is a resting iceberg order with $\mathit{visibleQty} = 0$ and
$\mathit{remainingQty} > 0$. It cannot be matched---$\mathit{fillQty} =
\min(\mathit{incoming.remainingQty}, 0) = 0$---and cannot reload because the
reload trigger is only in the fill path. The order is stranded: it occupies a
queue position, blocks no fills, and cannot be removed except by explicit
cancel.

\subsubsection*{Illustrative Scenario}

\begin{lstlisting}[language={},basicstyle=\ttfamily\footnotesize,frame=single,xleftmargin=0pt,framexleftmargin=0pt]
Resting iceberg: qty=2, displayQty=1, visibleQty=1,
                 remainingQty=2, stpGroup="G1"        (id=R)
Incoming:        qty=1, stpGroup="G1", stpPolicy=DECREMENT  (id=I)

DECREMENT:
    reduceQty = min(1, 1) = 1
    I.remainingQty = 0   -> I is cancelled
    R.visibleQty   = 0
    R.remainingQty = 1   -> R remains on book

Post-decrement state of R:
    visibleQty=0, remainingQty=1, displayQty=1
    -> Reload path not invoked
    -> R sits on book, invisible, unreachable by matching
\end{lstlisting}

\paragraph{Note on claim level.} This scenario is reachable in the prose
specification as written. However, the TLA+ model as checked incorporates the
fix (iceberg reload after DECREMENT) because the stranded state would cause
the model checker itself to loop: subsequent orders would find
$\mathit{visibleQty} = 0$ and skip the order without decrementing fuel,
creating an effective infinite loop in \texttt{DoMatch}. The bug is therefore
presented as a specification gap---a path the spec silently omits---rather
than as a TLC-witnessed invariant violation with a precise counterexample
trace.


\subsection{Design Observations (Not Demonstrated Bugs)}

The following interactions are noted as potential specification ambiguities.
They are not presented as verified bugs because no invariant violation was
produced for them.

\paragraph{Observation 1---FOK + STP DECREMENT.}
The FOK pre-check counts non-conflicting visible quantity. During matching,
DECREMENT can consume the incoming order's $\mathit{remainingQty}$ without
generating trades. The incoming order reaches $\mathit{remainingQty} = 0$ and
is treated as cancelled, but with fewer trade fills than $\mathit{qty}$.
Whether this constitutes a FOK violation depends on whether FOK is defined as
``fully traded'' or ``$\mathit{remainingQty}$ fully consumed.'' The spec does
not address this interaction.

\paragraph{Observation 2---MinQty + STP DECREMENT.}
Analogously, the MinQty pre-check guarantees
$\mathit{filled\_quantity} \geq \mathit{minQty}$ (\S5.4), but if DECREMENT
consumes part of the incoming quantity before any fills occur, the traded
quantity may fall below $\mathit{minQty}$ even though the pre-check passed.

\paragraph{Observation 3---MTL + MinQty clearing.}
The spec clears $\mathit{minQty}$ in two places: once in the MTL path (Phase
4) and once in the normal matching path (Phase 5a). This duplication is
currently correct but fragile---adding a new order type or routing path
requires remembering to add a third clearing point.


% ============================================================================
\section{Specification Fixes}
\label{sec:fixes}
% ============================================================================

Both confirmed issues were fixed in the TLA+ model. The fixes are minimal and
directly address the root causes.

\begin{table}[H]
\centering
\caption{Bug and fix summary.}
\label{tab:fixes}
\small
\begin{tabular}{@{}lL{3cm}L{3.5cm}L{4.5cm}l@{}}
\toprule
ID & Location & Description & Fix & Invariant \\
\midrule
BUG-1 & \S10.1 \texttt{EVALUATE\_STOPS}
      & Triggered stop retains original timestamp; violates FIFO when resting
      & Assign \texttt{stop.timestamp = currentTimestamp()} before re-entering pipeline
      & INV-7 \\
\addlinespace
BUG-2 & \S8.3 DECREMENT
      & Iceberg $\mathit{visibleQty}$ driven to 0 without reload; order stranded
      & After DECREMENT: if $\mathit{visibleQty} = 0 \land \mathit{remainingQty} > 0 \land \mathit{displayQty} \neq \bot$, reload visible slice, refresh timestamp, move to back
      & Spec gap \\
\bottomrule
\end{tabular}
\end{table}


\subsection{Fix for BUG-1 (Stop Timestamp)}

The fix is a one-line addition to \S10.1. In the TLA+ model,
\texttt{ProcessTriggeredStops} applies the fix:

\begin{lstlisting}[language=tlaplus]
converted == [ConvertStop(s) EXCEPT !.timestamp = tm]
\end{lstlisting}

The corresponding prose amendment to \S10.1:

\begin{lstlisting}[language={}]
-- After "Convert to base type":
stop.timestamp = currentTimestamp()   -- Assign trigger-time priority
PROCESS(stop, book)
\end{lstlisting}

This ensures that a triggered stop order receives a timestamp reflecting when
it entered the live matching pipeline, not when it was originally submitted.


\subsection{Fix for BUG-2 (Iceberg Reload After DECREMENT)}

The fix adds a post-DECREMENT iceberg check to \S8.3. In the TLA+ model:

\begin{lstlisting}[language=tlaplus]
ELSE IF newRest.visibleQty = 0 /\ newRest.displayQty /= NULL THEN
    LET reloadQty == Min(newRest.displayQty, newRest.remainingQty)
        reloaded  == [newRest EXCEPT !.visibleQty = reloadQty,
                                     !.timestamp = tm,
                                     !.status = "PARTIAL"]
        sc == SetContra(side, bQ, aQ, bestP,
                        restTail \o <<reloaded>>)
    IN DoMatch(newInc, sc.bQ, sc.aQ, tds, tm + 1, fuel - 1)
\end{lstlisting}

The corresponding prose amendment to \S8.3, after the DECREMENT quantity
update:

\begin{lstlisting}[language={}]
IF resting.visibleQty = 0 AND resting.remainingQty > 0
   AND resting.displayQty != null:
    resting.visibleQty = min(resting.displayQty, resting.remainingQty)
    resting.timestamp  = currentTimestamp()
    MOVE resting TO back OF level.orders
\end{lstlisting}

This mirrors the iceberg reload behavior specified in \S7.5 for the normal
fill path, extending it to the DECREMENT path for consistency.


% ============================================================================
\section{Lean~4 Mechanized Proof}
\label{sec:lean}
% ============================================================================

TLC searches a bounded state space and cannot establish correctness
for arbitrary book sizes. Production matching engines routinely hold
millions of orders, well beyond the configurations TLC can
enumerate. To lift the argument above that bound we translated the
fixed specification into a Lean~4 reference implementation and
mechanized a proof of the principal preservation property. This
section states the theorem, describes an early soundness issue with
a constant fuel bound, and presents two complementary proof
artifacts: a long constructive proof of the full preservation
theorem by branch-by-branch induction and a shorter structural proof
of the matching-and-dispose core by case analysis of the matching
termination condition.

Both proof files live in the same Lean project as the executable
matching engine and are type-checked on every build, with no open
obligations remaining. Table~\ref{tab:lean-stats} summarises their
sizes.

\begin{table}[H]
\centering
\caption{Lean~4 proof artifacts (final state).}
\label{tab:lean-stats}
\small
\begin{tabular}{@{}lrrl@{}}
\toprule
File & Lines & Theorems & Proof style \\
\midrule
\texttt{Theorems.lean}         & 3{,}986 & 74 & Constructive, branch-by-branch \\
\texttt{TheoremsElegant.lean}  &    312  &  6 & Three-case trichotomy \\
\midrule
\textbf{Total}                 & \textbf{4{,}298} & \textbf{80} & Complementary proof artifacts \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Goal and Statement}
\label{sec:lean-goal}

Define the combined book invariant
\[
\texttt{AllInv}(b) \;\equiv\;
\texttt{BookUncrossed}(b) \land
\texttt{bidsSortedDesc}(b) \land
\texttt{asksSortedAsc}(b).
\]
The principal result we establish is that \texttt{AllInv} is
preserved by the full processing pipeline under a minimal
well-formedness precondition. In Lean it is stated first as a mutual
induction over \texttt{processOrder}, \texttt{processCascade}, and
\texttt{processTriggeredStops}; the exported single-order theorem is:

\begin{lstlisting}[language={},basicstyle=\ttfamily\footnotesize]
theorem processOrder_preserves_AllInv
    (fuel : Nat) (o : Order) (b : BookState)
    (hpok : OrderProcOk o)
    (hstops : StopsNoPostOnly b)
    (h : AllInv b) :
    AllInv (processOrder fuel o b).book
\end{lstlisting}

The corresponding \texttt{process}-level \texttt{AllInv} corollary is
immediate because \texttt{process} only adds metadata updates
(\texttt{id}, \texttt{timestamp}, \texttt{nextId}, \texttt{clock})
and does not change bids or asks. We state the uncrossed-book
projection separately under the name
\texttt{process\_preserves\_uncrossed} because it is the property
most directly comparable to the TLC invariant
\texttt{BookUncrossed}, but the underlying proved result is the
stronger \texttt{AllInv} preservation statement.

The sortedness conjuncts of \texttt{AllInv} are not cosmetic.
\texttt{BookUncrossed} alone is not inductive with respect to the
matching loop: the proof does not close unless the induction
hypothesis also records that both sides of the book are monotone in
price. This was established by the induction refusing to close under
the weaker hypothesis, at which point the invariant was strengthened
to \texttt{AllInv} and the proof proceeded. We report this as an
example of the proof refining the invariant the specification should
have emphasized.

The preconditions \texttt{OrderProcOk} and \texttt{StopsNoPostOnly}
are not arbitrary hypotheses: they are minimal fragments of the
specification's own well-formedness rules (WF-13: post-only
$\Rightarrow$ LIMIT, and the derived LIMIT $\Rightarrow$ price
defined), also surfaced by the proof itself. We return to them in
\S\ref{sec:lean-wf}.

\subsection{The Fuel Problem}
\label{sec:lean-fuel}

Lean~4's termination checker requires every recursive function to be
provably total. The matching loop \texttt{doMatch} is recursive: one
step consumes a resting order, reduces the incoming order's quantity,
or removes a crossing price level, and recurses on the remainder.
Because the Lean termination checker cannot synthesize the measure
automatically, the original implementation used an explicit
\texttt{fuel~:~Nat} argument and a \emph{constant} bound:

\begin{lstlisting}[language={},basicstyle=\ttfamily\footnotesize]
def defaultFuel : Nat := 100
def doMatch (fuel : Nat) ... : MatchResult :=
  match fuel with
  | 0       => /* fuel exhausted: return as-is */
  | fuel'+1 => /* one step, recurse with fuel' */
\end{lstlisting}

This hides the assumption that matching always completes within 100
steps, which does not hold at arbitrary scale. Consider a book with
$200$ resting asks at price $50$ and one incoming BUY at price $100$.
\texttt{doMatch} consumes the first $100$ asks, exhausts its fuel,
and returns the incoming order with $100$ asks still resting at
price~$50$. The disposition phase then inserts the buy at price $100$
on the bid side, producing $\mathit{bestBid}=100 > \mathit{bestAsk}=50$---a
crossed book. The preservation theorem does not hold under the
constant-fuel assumption, and the proof fails to close.

This is the point at which the difference between the two
verification levels becomes concrete. TLC's explored configurations
never exceeded a handful of orders on either side, so the bounded
model-checking pass did not expose the issue; only the universally
quantified Lean statement forced the fuel bound to become a proof
obligation rather than an assumption. The arbitrary constant turned a
semantic property---``matching completes''---into a syntactic one---``matching
completes in $\leq 100$ steps''---and only the syntactic version was
provable.

\subsection{Computed Fuel and the Progress Lemma}
\label{sec:lean-fuel-fix}

The fix for the fuel unsoundness is to compute the fuel bound from the
book state and prove that the bound is always sufficient:

\begin{lstlisting}[language={},basicstyle=\ttfamily\footnotesize]
def matchMeasure (contra : List PriceLevel) (_inc : Order) : Nat :=
    totalRemaining contra + orderCount contra + contra.length

def computeMatchFuel (b : BookState) (side : Side) : Nat :=
    matchMeasure (contraLevels b side) dummyInc + 1
\end{lstlisting}

The measure has three components:
\texttt{totalRemaining}~$= \sum_{o \in \text{contra}}
o.\mathit{remainingQty}$ accounts for fills, \texttt{orderCount} accounts
for orders whose visible quantity is skipped without a fill (STP
cancellation, zero-visible iceberg slots), and \texttt{contra.length}
accounts for level exhaustion steps that remove an empty price level.

The key technical lemma---the \emph{progress lemma}---is that every
non-terminal step of \texttt{doMatch} strictly decreases this measure.
A single monolithic induction over all of \texttt{doMatch}'s branches
(there are 14) was intractable in our experience. The decomposition
that worked is per-branch: each branch becomes its own 10--25-line
lemma proving progress, and the 14 lemmas compose into a single
termination-and-stability theorem.
Table~\ref{tab:branches} summarizes the branches and the decrement each
one contributes.

\begin{table}[H]
\centering
\caption{Progress analysis for \texttt{doMatch} branches. Terminal
branches stop the loop; measure-decreasing branches recurse on a
strictly smaller state.}
\label{tab:branches}
\small
\begin{tabular}{@{}llL{5.5cm}@{}}
\toprule
Branch & Kind & Why the measure decreases \\
\midrule
Full fill (resting consumed)    & Decrease  & \texttt{orderCount} drops by 1 \\
Partial fill (incoming done)    & Terminal  & $\mathit{remainingQty}=0$ \\
STP cancel newest               & Terminal  & incoming cancelled \\
STP cancel oldest               & Decrease  & one order removed \\
STP cancel both                 & Terminal  & incoming cancelled \\
STP decrement                   & Decrease  & resting qty strictly reduced \\
Iceberg fill + reload           & Decrease  & fill reduced
                                              \texttt{totalRemaining} before reload \\
Level exhausted                 & Decrease  & empty level removed \\
Zero-visible skip               & Decrease  & one order removed \\
No crossable orders             & Terminal  & $\lnot\,\texttt{canMatchPrice}$ \\
\bottomrule
\end{tabular}
\end{table}

The iceberg reload case is the subtlest. Reload appears to ``add
liquidity back'': the resting order's visible quantity jumps from zero
to the display quantity. But \texttt{totalRemaining} is not affected by
this redistribution---only the visible/hidden split changes---and the
fill step that triggered the reload strictly decreased
\texttt{totalRemaining} by the filled quantity. The composition of the
two steps is a net decrease. The proof carries this through explicitly;
the reload is not a no-op, it is a book-keeping update that leaves the
measure strictly below its pre-step value.

With the progress lemma in place, the terminal states of
\texttt{doMatch} (on sufficient fuel) fall into a four-way stable
predicate: the incoming is consumed, the incoming is cancelled, the
contra side is empty, or the contra head does not cross the incoming
price. This predicate (\texttt{buyMatchStable}~/
\texttt{sellMatchStable}) is the black-box characterization that the
elegant proof in \S\ref{sec:lean-elegant} relies on.

\subsection{Side Parameterization}
\label{sec:lean-side}

Every matching lemma has both a buy-side and a sell-side form. The
first attempt duplicated every proof---\texttt{doMatch\_buy\_preserves\_bids}
alongside \texttt{doMatch\_sell\_preserves\_asks}, and so on for
fourteen branches on each side. The duplication doubled the proof
effort and made any change require two coordinated edits.

The fix is to parameterize every lemma by a \texttt{Side} value:

\begin{lstlisting}[language={},basicstyle=\ttfamily\footnotesize]
theorem doMatch_preserves_uncrossed
    (fuel : Nat) (inc : Order) (b : BookState)
    (tm : Timestamp) (side : Side)
    (hside : inc.side = side)
    (h : AllInv b) :
    BookUncrossed { b with
        bids := (doMatch fuel inc b.bids b.asks [] tm).bids,
        asks := (doMatch fuel inc b.bids b.asks [] tm).asks }
\end{lstlisting}

The proof then performs \texttt{cases side} \emph{only at the leaves}
where the concrete direction of the inequality matters. Everything
above the leaves is shared. After side-parameterization the sell mirror
of the fuel-sufficiency proof---which had been the bottleneck---compiled
on the first try.

\subsection{Mutual Induction Across the Cascade}
\label{sec:lean-mutual}

\texttt{processOrder}, \texttt{processCascade}, and
\texttt{processTriggeredStops} are mutually recursive: a trade can
trigger a dormant stop, which converts to a market order and re-enters
the pipeline, which can produce more trades, which can trigger more
stops. Proving \texttt{AllInv} preservation for any one of them
requires proving it for all three simultaneously:

\begin{lstlisting}[language={},basicstyle=\ttfamily\footnotesize]
theorem process_all_preserve_AllInv : forall fuel,
  (forall o b, OrderProcOk o -> StopsNoPostOnly b -> AllInv b ->
    AllInv (processOrder fuel o b).book) /\
  (forall ts b, StopsNoPostOnly b -> AllInv b ->
    AllInv (processCascade fuel ts b).book) /\
  (forall os b, (forall o in os, o.postOnly = false) ->
                StopsNoPostOnly b -> AllInv b ->
    AllInv (processTriggeredStops fuel os b).book)
\end{lstlisting}

Induction proceeds on \texttt{fuel}. The base case (\texttt{fuel = 0})
is trivial because all three functions return the book unchanged.
The step case must discharge every phase of \texttt{processOrder}:
stop handling, post-only rejection, FOK pre-check, MinQty pre-check,
market-to-limit routing, normal matching, dispose, and stop cascade.
Each phase follows the same template: apply
\texttt{doMatch\_preserves\_AllInv} to the matching step, extract the
non-crossing precondition from the termination-stability predicate,
apply \texttt{dispose\_preserves\_AllInv} to the disposition step, and
thread the cascade through the inductive hypothesis for
\texttt{processCascade}.

The subtlest sub-case is \emph{MTL Phase~4 case~3}: a MARKET-TO-LIMIT
order that fills partially, converts to a LIMIT order, then runs a
\emph{second} matching pass with the derived limit price, then
disposes. The second \texttt{doMatch} operates on the post-first-match
book with a freshly constructed incoming order, and the non-crossing
precondition for the subsequent dispose must be extracted from a
\texttt{doMatch} call that does not match the form of any of the
existing helper lemmas. We resolved this by parameterizing the
non-crossing extraction helpers by the starting timestamp
(\texttt{matching\_dispose\_noCross\_buy}~/~\texttt{sell} each take an
explicit \texttt{tm}) so they apply uniformly to both the first
matching pass (which uses \texttt{b.clock + 1}) and the MTL second pass
(which uses \texttt{mr.clock}).

\subsection{Minimal Well-Formedness Preconditions}
\label{sec:lean-wf}

The precondition \texttt{OrderProcOk~o} captures two facts that the
proof cannot avoid:

\begin{lstlisting}[language={},basicstyle=\ttfamily\footnotesize]
def OrderProcOk (o : Order) : Prop :=
    (o.postOnly = true -> o.price.isSome = true) /\
    ((o.orderType = .stopLimit \/ o.orderType = .stopMarket) ->
       o.postOnly = false)
\end{lstlisting}

The first conjunct rules out the post-only / no-price configuration
that the spec forbids via WF-13 (\emph{post-only $\Rightarrow$
LIMIT}) and WF-2 (\emph{LIMIT $\Rightarrow$ price defined}). In the
proof, this hypothesis discharges the otherwise-unreachable branch in
which \texttt{wouldCross} returns false for a post-only order with
\texttt{price = none}. Without it, the Phase-2 case of
\texttt{processOrder} has a genuinely crossed output.

The second conjunct---stop orders are not post-only---is a
consequence of the same WF-13 by contrapositive: post-only
$\Rightarrow$ LIMIT, and stop types are not LIMIT. It is needed
when the stop-cascade arm of \texttt{processOrder} converts a
triggered stop and re-enters the pipeline.

A separate precondition,
\texttt{StopsNoPostOnly~b}~$\equiv~\forall s \in
b.\mathit{stops},~s.\mathit{postOnly} = \texttt{false}$, lifts the
same constraint to the resting stop pool. It is preserved by every
pipeline phase and is used by the stop-cascade case.

Both preconditions were \emph{discovered}, not guessed. We first
attempted the proof without them; the proof failed at a specific
branch with a concrete counterexample; the counterexample corresponded
exactly to a state that the spec's own WF rules forbid. The fragment
we added is strictly weaker than the full \texttt{Order.wellFormed}
predicate---we only require what the proof actually needs. In
particular, we do \emph{not} require the incoming order to pass the
full twenty-constraint WF check; only the two clauses above. This is
the honest precondition: the proof tells us what matters.

\subsection{The Elegant Alternative Proof}
\label{sec:lean-elegant}

Once the constructive proof was complete, we developed a second
proof component, in the file \texttt{TheoremsElegant.lean}
(6~theorems, 312~lines), that recasts the matching-and-dispose
core as a shorter structural argument.

When \texttt{doMatch} terminates on a fuel-sufficient run, the
resulting state falls into one of three cases:

\begin{enumerate}[nosep]
\item \textbf{Aggressor consumed.} $\mathit{remainingQty} = 0$ (or the
      order has been cancelled). The incoming order is gone;
      \texttt{doMatch} only removes orders from the contra side, never
      adds to it; removing orders cannot create a cross.
\item \textbf{Contra side empty.} There is nothing left on the
      opposite side to cross against; an empty side is vacuously
      non-crossing.
\item \textbf{Non-crossing head.} Both sides are non-empty but the
      contra head does \emph{not} cross the incoming price. This is
      the only way \texttt{doMatch} can stop with both sides active:
      if prices crossed, the loop guard \texttt{canMatchPrice} would
      fire and matching would take another step. Since it stopped,
      prices do not cross.
\end{enumerate}

A fourth case---both sides active and still crossing---is not
reachable at termination: such a state is not terminal under the
loop guard, so fuel sufficiency rules it out. The three cases
compose with \texttt{dispose} directly: Case~1 returns the book
unchanged, Case~2 inserts on a book whose contra side is empty
(vacuously non-crossing), and Case~3 inserts at a price strictly
separated from the contra head.

The second proof treats \texttt{doMatch} as a black box
characterised by its termination condition rather than by the
control-flow path that produced it. It reuses the constructive
proof's fuel-sufficiency lemma (\texttt{doMatch\_output\_stable}) as
an imported fact.

\begin{center}
\small
\begin{tabular}{@{}lll@{}}
\toprule
                        & Constructive              & Structural              \\
\midrule
File                    & \texttt{Theorems.lean}    & \texttt{TheoremsElegant.lean} \\
Lines                   & 3{,}986                   & 312                     \\
Theorems/lemmas         & 74                        & 6                       \\
Argument                & Branch-by-branch          & Three-case analysis     \\
\bottomrule
\end{tabular}
\end{center}

Both are machine-checked, and both live in the build. The
constructive proof discharges fuel sufficiency, sortedness
preservation, side isolation, and the mutual pipeline theorem. The
structural file reuses those lemmas and shows that, once the
termination condition is fixed, the matching-and-dispose argument
itself is short.

\subsection{Model Checking and Theorem Proving Are Complementary}
\label{sec:lean-complementary}

TLC and Lean surfaced different classes of issue. TLC produced the
two feature-composition defects (FIFO via stale timestamps; iceberg
stranding under STP DECREMENT) as concrete multi-step traces that a
proof attempt would not have generated directly. Lean's universal
quantification, in turn, made the constant-fuel assumption
(\S\ref{sec:lean-fuel}) visible; TLC did not reach scopes where
the fuel bound mattered.

At the level of cost, TLC returned its first counterexample in 102
seconds of wall time, while the Lean proof accumulated weeks of
interactive development and produced a 4{,}298-line artifact. The
two are complementary rather than substitutable: TLC's speed is
useful early, and the Lean statement scales beyond any bounded
configuration.


% ============================================================================
\section{Specification-Guided Implementation}
\label{sec:synthesis}
% ============================================================================

For completeness we describe how the fixed specification was carried
forward into systems-language implementations. We frame this as an
engineering observation, not as an additional verification result:
we do not prove refinement between the Lean reference and the C or
C++ code, and the assurance we obtain at the implementation level is
bounded by shared tests and runtime invariant checks rather than by
a mechanized proof.

Prose and TLA+ artifacts from \S\ref{sec:semantics}--\S\ref{sec:fixes}
were used as the input to a language-model-assisted coding workflow,
augmented with two architectural directives: intrusive doubly-linked
lists for $O(1)$ order removal and FIFO management, and a pool
allocator to bound allocation latency. The resulting C and C++
implementations follow the phase structure of the specification one
for one, and apply the two defect fixes introduced in \S\ref{sec:fixes}:
stop-trigger timestamps are refreshed before re-entry into the
matching pipeline, and the STP DECREMENT path reloads iceberg
visible quantity when it would otherwise reach zero.

The C implementation additionally includes an
\texttt{engine\_check\_invariants} routine that translates the
TLC-checked invariants into runtime assertions, giving a form of
continuous conformance checking between the running engine and the
invariant suite. This is weaker than a refinement proof---an
invariant assertion fires after state has already diverged---but it
narrows the gap between specification and implementation and has
caught regressions during development.

We deliberately avoid framing this step as ``correct-by-construction''
code generation. A precise specification is an unusually good input
for an implementation effort, and a feature-rich matching engine is
an unusually tractable target because its control-flow structure is
exposed by the specification itself. The two defect fixes were
carried across because the specification was explicit about them,
not because the generation step carried any inherent guarantee. A
genuine end-to-end guarantee would require a refinement argument
connecting the Lean reference to the systems-language code, and that
is beyond the scope of this paper.


% ============================================================================
\section{Lessons Learned}
\label{sec:lessons}
% ============================================================================

\subsection{Feature Composition Is Where Bugs Hide}

Both findings share a structural pattern: each involved feature is correctly
specified in isolation, but the specification omits a clause covering their
interaction. Stop orders are correctly described; iceberg reload is correctly
described; FIFO insertion is correctly described. None of these sections
contains an error when read alone. The bug lives in the intersection: what
happens to an iceberg order's timestamp when the mechanism that zeros its
visible quantity is not the fill path but the STP DECREMENT path?

This is not a failure of careful writing---it is a structural limitation of
prose specifications. Prose sections have natural scope boundaries.
Interactions between features that span sections are systematically harder to
see than behaviors within a single section.

A formal model collapses all features into a single operational semantics and
evaluates invariants globally. The intersection cases are not special---they
are just more states.


\subsection{Logical Clocks Amplify Ordering Bugs}

The FIFO invariant is stated in terms of timestamps. Because the
specification uses a monotonically increasing logical clock, any state
transition that fails to advance the clock before inserting an order can
produce a stale timestamp that violates ordering. The stop-trigger bug is
exactly this: the clock was not advanced before the triggered stop was assigned
to the book.

Timestamp-based ordering invariants are fragile in the presence of features
(like stop triggers) that re-enter orders into the pipeline from a dormant
pool. Wherever an order exits a dormant state and re-enters an active queue,
its timestamp must be refreshed. The fix is simple; the risk is that prose
specifications do not enumerate every re-entry point.


\subsection{Iceberg Reload Is Not Tied Tightly Enough to Its Trigger Condition}

The iceberg reload semantics in \S7.5 are described in terms of the fill path:
``when $\mathit{visibleQty}$ is fully consumed [by a fill].'' The prose
implicitly assumes that $\mathit{visibleQty}$ can only reach zero through the
fill path. STP DECREMENT provides a second path to zero
$\mathit{visibleQty}$ that the reload trigger does not anticipate.

A more robust specification would state the reload condition declaratively as
an invariant on the resting order state:

\begin{quote}
A resting order must not have $\mathit{visibleQty} = 0$ and
$\mathit{remainingQty} > 0$.
\end{quote}

\noindent Rather than specifying reload as a step triggered by the fill path,
it would be specified as a consequence of this invariant: any action that could
produce this state must be followed by a reload. This declarative framing
would make the omission in DECREMENT immediately apparent.


\subsection{The Specification Artifact Has Independent Value}

The TLA+ model serves as a precise, executable reference for the prose
specification. Independently of the bugs found, it forces every semantic
choice to be made explicit: what is the recursion structure of stop cascades?
What exactly does ``timestamp'' mean for a triggered order? At what point
does MTL price derivation occur relative to the cascade? These questions have
definite answers in the model even when the prose is ambiguous.


% ============================================================================
\section{Limitations and Reproducibility}
\label{sec:limitations}
% ============================================================================

\subsection{Bounded Exhaustive Exploration, Not Proof (for TLA+)}

TLC performs bounded exhaustive state enumeration: it finds every state
reachable from \texttt{Init} via the \texttt{Next} relation, within the
parameter bounds. A clean run means no invariant violation was found within
the explored configurations. It does not mean no violation exists for larger
configurations, more orders, or more price levels. We do not claim the
TLA+ model is correct in any unbounded sense.

Specifically:
\begin{itemize}[nosep]
\item The 3-order partial run did not complete state-space exploration.
\item No configuration with more than two distinct STP groups was checked.
\item GTD time-based expiry was checked only via the simplified
      \texttt{TimeAdvance} action, not against wall-clock constraints.
\item The event-ordering invariant (INV-10) is not formalized in the model.
\end{itemize}

\paragraph{Lean closes the unbounded gap for \texttt{AllInv}, not
for the full invariant suite.} The Lean proof in
\S\ref{sec:lean} establishes \texttt{process\_preserves\_AllInv}
for arbitrary book sizes and arbitrary order counts, under the
minimal preconditions \texttt{OrderProcOk} and
\texttt{StopsNoPostOnly}. The corollary
\texttt{process\_preserves\_uncrossed} is an immediate consequence.
Among the broader TLC-checked invariants, only the \texttt{AllInv}
portion (INV-2, INV-3, INV-4) has been lifted to Lean. The
remaining invariants---no ghosts (INV-5), status consistency
(INV-6), FIFO within level (INV-7), no resting markets
(INV-8), post-only safety (INV-11), STP guarantee (INV-12), no
resting MTL (INV-13), no resting MinQty (INV-14)---remain at the
bounded level of assurance provided by TLC. Lifting them to Lean is
future work.


\subsection{Specification vs.\ Implementation}

\textbf{For the Lean reference implementation, the gap is closed.} The
Lean~4 code that \texttt{\#eval} runs in tests is the same code that
the theorems reference---there is no separate model, no extraction
layer, and no refinement proof obligation. For the Lean artifact, the
semantic gap between tested behavior and proved behavior is zero.

\textbf{For the C and C++ synthesized implementations, a gap
remains.} The Lean proof does not transfer automatically to C or C++;
those implementations are validated by shared test suites derived from
the specification, not by a formal correspondence argument. An
implementation that correctly implements the (fixed) specification
could still contain implementation bugs---off-by-one errors,
concurrency hazards, data-structure corruption---that are outside the
scope of this work. A refinement proof connecting the Lean reference
to the systems-language implementations would close the remaining gap
and is a natural next step.


\subsection{Reproducibility}

All artifacts described in this paper---the prose specification,
the TLA+ model and configurations, the Lean~4 project, and the
paper source---are available in a public Git repository at
\url{https://github.com/formally-verified-exchange/verified-matching-engine}.
The top-level \texttt{README.md} lists prerequisites and gives
step-by-step commands for building the Lean proofs and running
TLC against each configuration. The complete artifacts are:
\begin{itemize}[nosep]
\item \texttt{matching-engine-formal-spec.md}---the prose specification.
\item \texttt{spec/MatchingEngine.tla}---the TLA+ specification
      ($\sim$950 lines).
\item \texttt{spec/MatchingEngine.cfg} and the per-configuration files
      \texttt{spec/MatchingEngine\_*.cfg} for Table~\ref{tab:configs}.
\item \texttt{spec/REPORT.md}---model checking output and bug descriptions.
\item \texttt{MatchingEngine/}---the Lean~4 project (11 source
      files). The two proof files are
      \texttt{Theorems.lean} (3{,}986 lines, 74 theorems) and
      \texttt{TheoremsElegant.lean} (312 lines, 6 theorems), both
      under the \texttt{MatchingEngine/} directory. Both are built
      and type-checked by \texttt{lake build} and are imported from
      the library root \texttt{MatchingEngine.lean}.
\item Synthesized C/C++ implementations (available in related
      repositories).
\end{itemize}

To reproduce the FIFO counterexample:
\begin{enumerate}[nosep]
\item Revert the stop timestamp fix in \texttt{ProcessTriggeredStops}: replace\\
      \texttt{[ConvertStop(s) EXCEPT !.timestamp = tm]} with
      \texttt{ConvertStop(s)}.
\item Run TLC with the Small configuration
      (\texttt{MAX\_ORDERS=2, MAX\_QTY=2, PRICES=\{1,2\}}, amend disabled).
\item TLC will find a violation of \texttt{FIFOWithinLevel} within the first
      several million states.
\end{enumerate}

To reproduce the fuel-unsoundness counterexample (\S\ref{sec:lean-fuel}):
\begin{enumerate}[nosep]
\item Revert \texttt{computeMatchFuel} in
      \texttt{MatchingEngine/Process.lean} to the constant
      \texttt{defaultFuel~:=~100}.
\item Re-run \texttt{lake build}.  The proof of
      \texttt{doMatch\_preserves\_AllInv} will fail: the measure
      argument no longer dominates the recursive call bound, and any
      book with more than $100$ contra-side orders produces a
      concrete cross after one application of \texttt{process}.
\end{enumerate}


% ============================================================================
\section{Related Work}
\label{sec:related}
% ============================================================================

The work we know of falls into three largely disjoint literatures:
formal specification and model checking of distributed systems,
mechanized proof of systems software, and economic analysis of
exchange microstructure. Each addresses part of what this paper
covers, but none combines a feature-rich limit-order-book
specification with both bounded bug-finding and mechanized
theorem-level assurance. The positioning below maps the closest
neighbors.

\paragraph{TLA+ in industry.}
TLA+ has been applied at scale at Amazon Web
Services~\cite{Newcombe2015}, where engineers used it to model DynamoDB,
S3, and other distributed services and uncovered design-level bugs
that would have been difficult to find through testing. Similar
reports have come from Microsoft (Azure Cosmos DB), MongoDB, and
Elastic. Our work follows the Newcombe et al.\ template---use TLC to
explore a specification's state space and find interaction bugs
before coding---but differs in the target domain (a financial
matching engine rather than a distributed storage protocol) and in
the follow-through to a machine-checked proof in a different tool
(Lean~4) rather than stopping at bounded model checking.

\paragraph{Mechanized proof of systems software.}
The largest-scale machine-checked software artifacts---the seL4
microkernel~\cite{Klein2009}, CompCert~\cite{Leroy2009}, and
CertiKOS~\cite{Gu2016}---prove functional correctness or refinement
between a high-level specification and a C-level implementation,
and consumed person-decades of effort. The present work is
considerably narrower in both ambition and cost: we prove
preservation of a single book invariant (\texttt{AllInv}) for the
Lean reference implementation of a feature-rich matching engine
and rely on a separate testing-and-assertion path to connect that
reference to systems-language code. The investment is
correspondingly smaller (weeks rather than years), but the target
is a different class of system---one whose complexity lives in
feature composition rather than in concurrency or low-level memory
semantics.

\paragraph{Formal methods for financial systems.}
Prior formal work on financial infrastructure has focused mostly on
auction-theoretic properties of matching
algorithms~\cite{Budish2015} and on smart-contract verification
for decentralized exchanges~\cite{Bhargavan2016,Grishchenko2018}.
Property-based testing tools like QuickCheck~\cite{Claessen2000} are
used at some trading firms for engine validation, but published
accounts of end-to-end formal verification of a production-style
limit-order-book engine with the full order-type suite (icebergs,
stops, STP, post-only, FOK, MinQty) are, to our knowledge, absent
from the literature. A TLA+ specification of an order book has been
circulated informally in the TLA+ user community, but with a
substantially smaller feature set and no mechanized proof layer.

\paragraph{Termination proofs and well-founded recursion.}
The core technical device in \S\ref{sec:lean-fuel-fix}---replacing an
arbitrary fuel bound with a state-derived measure and a progress
lemma---is a standard pattern in mechanized proofs of imperative
programs~\cite{Manolios2006,Leino2013}. Our contribution is not the
pattern itself but a concrete demonstration that an LLM-assisted
proof can be pushed through the per-branch progress argument for a
matching-loop implementation with 14 interacting branches, closing
the bug introduced by the arbitrary constant.

\paragraph{LLM-assisted formal verification.}
A growing body of work explores the use of large language models to
draft Lean, Coq, or Isabelle tactics~\cite{Yang2023,First2023}.
These systems focus on automating tactic search. Our experience is
complementary: the AI did not autonomously close the proof through
tactic search; it wrote naturalistic Lean proofs under human
strategic direction. The progress-lemma decomposition, the
side-parameterization refactor, and the three-case trichotomy were
human contributions. The ~4{,}000 lines of case-splitting and
omega invocations were the AI's contribution. Neither layer alone
would have produced the artifact.

\paragraph{Spec-to-implementation synthesis.}
Program synthesis from formal specifications is a classical research
area~\cite{Manna1980}. Recent work on LLM-driven
synthesis~\cite{Jain2022} has reached practical levels for small
programs. Our ``Guided Synthesis'' paradigm differs in that the
specification is both the logical source of truth \emph{and} the
prompt, and architectural hints (intrusive lists, pool allocation)
are provided as orthogonal performance directives. The result is
then validated for conformance via shared test vectors derived from
the spec. A formal refinement argument between the Lean reference
and the synthesized C/C++ is future work.


% ============================================================================
\section{Conclusion}
\label{sec:conclusion}
% ============================================================================

We formalized a feature-rich matching-engine specification in TLA+
and ran bounded exhaustive exploration across several configurations
totaling over 36~million distinct states. TLC identified one
specification defect---a FIFO violation caused by a triggered stop
order retaining its original submission timestamp---and one
specification gap in the interaction between STP DECREMENT and
iceberg visibility. The FIFO counterexample is a three-order
sequence whose minimality is part of the finding: reading the stop,
insertion, and FIFO sections of the prose specification individually
does not expose it, but an operational model that composes all three
does.

We then proved, in Lean~4, that the processing pipeline preserves
the combined book invariant \texttt{AllInv} (uncrossed plus sorted
on both sides) under a minimal well-formedness precondition, for
arbitrary book sizes. The matching-loop fuel bound is derived from
the book state rather than assumed as a constant.
\texttt{Theorems.lean} (3{,}986 lines, 74 theorems) proves the full
preservation theorem branch-by-branch through the 14 branches of the
matching loop and the 11 phases of \texttt{processOrder}.
\texttt{TheoremsElegant.lean} (312 lines, 6 theorems) adds a shorter
structural argument for the matching-and-dispose core via a
three-case analysis of the matching termination condition. Both
files are type-checked by every build and contain no open
obligations.

One issue surfaced during the proof effort itself: the original
arbitrary fuel bound made the preservation statement false for
sufficiently populated books. Replacing it with a state-derived
measure and a progress lemma restores soundness.

The broader takeaway is that bounded model checking and mechanized
theorem proving play complementary roles on the same specification.
TLC is effective at surfacing concrete feature-composition defects
inside small state bounds and at speeds useful during early
specification work. Lean provides guarantees that hold for arbitrary
scale and, as a side effect, exposes pipeline issues that bounded
exploration cannot reach. Neither tool substitutes for the other.


\section*{Acknowledgments}

The TLA+ model, the Lean~4 proofs, and the C/C++ implementations
discussed in \S\ref{sec:synthesis} were developed interactively with
the assistance of a large language model, using the formal
specification as the primary input. The constant-fuel soundness
issue described in \S\ref{sec:lean} was encountered during this
process.

% ============================================================================
\begin{thebibliography}{99}
% ============================================================================

\bibitem{Newcombe2015}
C.~Newcombe, T.~Rath, F.~Zhang, B.~Munteanu, M.~Brooker, and M.~Deardeuff.
How Amazon Web Services Uses Formal Methods.
\emph{Commun. ACM}, 58(4):66--73, 2015.

\bibitem{Klein2009}
G.~Klein et al.
seL4: Formal Verification of an OS Kernel.
In \emph{Proc. ACM SOSP}, 2009.

\bibitem{Leroy2009}
X.~Leroy.
Formal Verification of a Realistic Compiler.
\emph{Commun. ACM}, 52(7):107--115, 2009.

\bibitem{Gu2016}
R.~Gu et al.
CertiKOS: An Extensible Architecture for Building Certified
Concurrent OS Kernels.
In \emph{Proc. USENIX OSDI}, 2016.

\bibitem{Budish2015}
E.~Budish, P.~Cramton, and J.~Shim.
The High-Frequency Trading Arms Race: Frequent Batch Auctions as a
Market Design Response.
\emph{Quarterly Journal of Economics}, 130(4):1547--1621, 2015.

\bibitem{Bhargavan2016}
K.~Bhargavan et al.
Formal Verification of Smart Contracts: Short Paper.
In \emph{Proc. PLAS}, 2016.

\bibitem{Grishchenko2018}
I.~Grishchenko, M.~Maffei, and C.~Schneidewind.
A Semantic Framework for the Security Analysis of Ethereum Smart
Contracts.
In \emph{Proc. POST}, 2018.

\bibitem{Claessen2000}
K.~Claessen and J.~Hughes.
QuickCheck: A Lightweight Tool for Random Testing of Haskell Programs.
In \emph{Proc. ACM ICFP}, 2000.

\bibitem{Manolios2006}
P.~Manolios and D.~Vroon.
Termination Analysis with Calling Context Graphs.
In \emph{Proc. CAV}, 2006.

\bibitem{Leino2013}
K.~R.~M. Leino.
Developing Verified Programs with Dafny.
In \emph{Proc. ICSE}, 2013.

\bibitem{Yang2023}
K.~Yang et al.
LeanDojo: Theorem Proving with Retrieval-Augmented Language Models.
In \emph{Proc. NeurIPS}, 2023.

\bibitem{First2023}
E.~First, M.~Rabe, T.~Ringer, and Y.~Brun.
Baldur: Whole-Proof Generation and Repair with Large Language Models.
In \emph{Proc. ESEC/FSE}, 2023.

\bibitem{Manna1980}
Z.~Manna and R.~Waldinger.
A Deductive Approach to Program Synthesis.
\emph{ACM TOPLAS}, 2(1):90--121, 1980.

\bibitem{Jain2022}
N.~Jain et al.
Jigsaw: Large Language Models Meet Program Synthesis.
In \emph{Proc. ICSE}, 2022.

\end{thebibliography}

\end{document}
